---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>

# MVP

1. Load the `diamonds.csv` data set and undertake an initial exploration of the data. You will find a description of the meanings of the variables on the relevant [Kaggle page](https://www.kaggle.com/shivam2503/diamonds/)

<br>

```{r}
library(tidyverse)
library(GGally)
library(janitor)
library(fastDummies)
library(ggfortify)
```

```{r}
diamonds <- clean_names(read_csv("data/diamonds.csv"))
glimpse(diamonds)
```


2. We expect the `carat` of the diamonds to be strong correlated with the physical dimensions `x`, `y` and `z`. Use `ggpairs()` to investigate correlations between these four variables.

<br>
```{r}
diamonds_trimmed <- diamonds %>% 
  select(c("carat", "x", "y", "z"))
ggpairs(diamonds_trimmed)
```

3. So, we do find significant correlations. Let's drop columns `x`, `y` and `z` from the dataset, in preparation to use only `carat` going forward.

<br>
```{r}
diamonds_new <- diamonds %>% 
  select(-c("x", "y", "z"))
```

4. We are interested in developing a regression model for the `price` of a diamond in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `price` and the predictors (this may take a while to run, don't worry, make coffee or something).

  ii. Perform further `ggplot` visualisations of any significant correlations you find.
  
<br>
```{r}
ggpairs(diamonds_new)
```

```{r}
diamonds_new  %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_point()
```

5. Shortly we may try a regression fit using one or more of the categorical predictors `cut`, `clarity` and `color`, so let's investigate these predictors: 

  i. Investigate the factor levels of these predictors. How many dummy variables do you expect for each of them?
cut: there are 5 levels, so 4 dummy variables are required
```{r}
diamonds_new %>% 
  distinct(cut)
```


clarity: there are 8 levels, so 7 dummy variables are required
```{r}
diamonds_new %>% 
  distinct(clarity)
```

colour: there are 7 levels, so 6 dummy variables are required
```{r}
diamonds_new %>% 
  distinct(color)
```

  ii. Use the `dummy_cols()` function in the `fastDummies` package to generate dummies for these predictors and check the number of dummies in each case.
cut: 
```{r}
diamonds_new %>% 
  dummy_cols(select_columns = "cut",
             remove_first_dummy  = TRUE)
```

clarity
```{r}
diamonds_new %>% 
  dummy_cols(select_columns = "clarity",
             remove_first_dummy = TRUE)
```

colour
```{r}
diamonds_new %>% 
  dummy_cols(select_columns = "color",
             remove_first_dummy = TRUE)
```

    
6. Start with simple linear regression. Regress `price` on `carat` and check the regression diagnostics.

<br>
```{r}
diamonds_new %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
diamonds_linear_model <- lm(price ~ carat, data = diamonds_new)

autoplot(diamonds_linear_model)
summary(diamonds_linear_model)
```


7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.


    
<hr>
```{r}
diamonds_multiregression_model1 <- lm(price ~ carat + table, data = diamonds_new)
autoplot(diamonds_multiregression_model1) # the output does not look very promising
summary(diamonds_multiregression_model1) # doesn't add much explanatory power - r2 went up only about 1pp
```


# Extension
    
8. Try adding an interaction between `log(carat)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction
    