---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

```{r}
library(tidyverse)
library(janitor)
library(factoextra)
```

```{r}
customers <- clean_names(read_csv("data/mall_customers.csv"))
```

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

```{r}
customers_trimmed <- customers %>%
  mutate_if(is.numeric, scale) %>% 
  mutate(gender = ifelse(gender == "Male",1,0)) %>% 
  select(-customer_id)
```

```{r}
# scaling the data
customers_scaled <- customers_trimmed %>% 
  mutate_if(is.numeric, scale)
```

```{r}
cluster_cust <- kmeans(customers_scaled, 
                      centers = 6,
                      nstart = 25)
cluster_cust
```

* Perform k-means clustering and chose a value of k.
```{r}
library(broom)

tidy(cluster_cust,
     col.names = colnames(customers_scaled))

augment(cluster_cust, customers)

```
```{r}
glance(cluster_cust)
```
```{r}
# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(customers_scaled, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers)
  )

k_clusters %>% 
  unnest(augmented)
```

Choosing your number of clusters
```{r}
clusterings <- k_clusters %>% 
  unnest(glanced)

clusterings

ggplot(clusterings) +
  aes(x = k, y = tot.withinss) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1,20, by = 1))
```

Silhouette Coefficient
>0 means you data is well clustered (the closer to 1 the better)
= 0 means it's close to a decision boundary
<0 means it was placed in the wrong cluster

average SC over all clusters.  average over 0.5 is good
The value of k (number of clusters) that gives you the best SC.

```{r}
fviz_nbclust(customers_scaled,
             kmeans,
             method = "silhouette",
             nstart = 25)
```

Gap statistic
```{r}
fviz_nbclust(customers_scaled,
             kmeans,
             method = "gap_stat",
             nstart = 25)
```

```{r}
names(customers_trimmed)
```


* Visualise the clustering for your chosen value of k.
```{r}
clusterings %>% 
  unnest(cols = c(augmented)) %>% 
  filter(k == 2) %>% 
  ggplot() +
  aes(x = age, y = spending_score_1_100, colour = .cluster, label = gender) +
  geom_point(aes(colour = .cluster)) + 
  geom_text(hjust = 0, vjust = -0.5, size = 3)
```

* Do you think the clustering seems a good fit for this data?
Yes, because there does seem to a be 2 distinct clusters on age and spending.

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 






