---
title: "Titanic homework with probability"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)

titanic <- read_csv("data/titanic.csv")

unique(titanic$class)
```
```{r}
# create a `crew` column containing values `TRUE` or `FALSE`
crew <- c("engineering crew", "victualling crew", "restaurant staff", "deck crew")
titanic <- titanic %>% 
  mutate(crew = if_else(class %in% crew, TRUE, FALSE))
  
```

```{r}
titanic %>%
  count(crew)
```


prob( survived ∣ crew )  
```{r}
p_survived_crew  <- titanic %>% 
  filter(survived == "yes", crew == TRUE) %>% 
    summarise(p = n()/nrow(titanic))
p_survived_crew
```


prob( survived ∣ passenger ) 
```{r}
p_survived_passenger  <- titanic %>% 
  filter(survived == "yes", crew == FALSE) %>% 
    summarise(p = n()/nrow(titanic))
p_survived_passenger
```

If I roll a six-sided dice once, what is the probability of getting a number less than 5?
```{r}
s <- rolldie(1)
(size_s <- nrow(s))

less_than_five <- s %>% 
  dplyr::filter(X1  < 5)
(size_less_than_five <- nrow(less_than_five))

prob_less_than_five <- size_less_than_five / size_s
prob_less_than_five
```

2. How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
6^3
```
216 outcomes in the sample set

4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint This is an arrangement where order matters. Of all the possible arrangements, how many are correct?
```{r}
people <- c("A", "B", "C", "D", "E")

chair_permu <- as_tibble(
  permutations(
    n = 5,
    r = 5,
    v = people,
    repeats.allowed = FALSE)
)

nrow(chair_permu)

1/nrow(chair_permu)
```

Only one correct seating is possible.
The probability of this outcome is 1/120

Other solution:
```{r}
# number of ordered arrangements:
num_outcomes <- factorial(5)
# only one of these is correctly ordered:
prob_perfect_order <- 1/num_outcomes
prob_perfect_order
```


6. If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

Hints

Investigate the cards() function in the prob package, if you’d like to try this in R (make a deck without jokers).
You could think about adding your own black_card random variable to the sample space, which would be TRUE if the suit is ‘Spade’ or ‘Club’, and FALSE otherwise.
```{r}
cards()
```

