---
title: "Homework - CIs"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>

# MVP

Now we'll go back to CI creation in the normal fashion. We'll take the `ames` data from the CIs lab earlier today and regard it now as a **sample**, we won't be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!

<br>

**Task 1.**  
Load the data again, `clean_names()`, and re-familiarise yourself with it

<br>
```{r}
library(tidyverse)
library(infer)
library(janitor)

ames <- read_csv("data/ames.csv") %>% 
  clean_names()
```


**Task 2.**  
Investigate the distribution of `lot_area`. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot(aes(lot_area)) +
  geom_histogram(col = "white", bins = 100)
```
looks broadly normally-shaped

<br>

**Task 3.**  
Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.

<br>
```{r}
infer_resample <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

infer_resample
```
```{r}
infer_resample %>% 
  ggplot(aes(stat)) +
  geom_histogram(col = "white", bins = 100)
```


**Task 4.**  
Use your bootstrap distribution to calculate a $95\%$ CI for `mean(lot_area)`, and visualise it on the distribution

<br>
```{r}
infer_ci_95 <- infer_resample %>% 
  get_ci(level = 0.95, type = "percentile")

infer_ci_95
```

```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci_95)
```
**Task 5.**  
You would like to know the `mean(lot_area)` of the sold houses with higher confidence. Calculate the $99\%$ CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the $95\%$ CI? Does that make sense?

<br>
```{r}
infer_ci_99 <- infer_resample %>% 
  get_ci(level = 0.99, type = "percentile")

infer_ci_99
```
The 99% interval is broader than the 95% interval, which makes sense.

**Task 6.**  
Calculate the point estimate of the `mean(lot_area)`

<hr>
```{r}
infer_resample %>%   
  summarise(mean = mean(stat))
```
The point estimate of the mean(lot_area) is 10,147 (square feet).

# Extension

<br>

**Task 1.**  
Calculate a point estimate and $95\%$ CI for the proportion of houses in the data built before 1920.  Does the number of `reps` you use matter? [Investigate `reps` from $200$ up to $50000$, memory of your laptop permitting].
<br><br>

```{r}
old_houses <- ames %>% 
  filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% 
  generate(reps = 100, type = "bootstrap") %>% 
  calculate(stat = "mean") 
old_houses
```

```{r}
old_house_mean <- old_houses%>% 
  summarise(mean = mean(stat)) 
old_house_mean
```


```{r}
old_house_ci_99 <- old_houses %>% 
  get_ci(level = 0.99, type = "percentile")

old_house_ci_99
```

